---
title: "HW_2"
author: "Erin Anderson"
date: "2026-09-27"
output: html_document
---
## Question 1

1. The data for this problem is on the speed (X, in mph) and stopping distance (Y, in ft) of 50
cars. What would be the power of an α=.05 test if the true slope of the regression line for
this problem is 1.7? Assume σ{b1} = 0.5. Write a sentence explaining what this power
means.

```{r setup, include=TRUE}
# delta = abs(beta1 - beta(10))/sigma(beta1)

delta <- abs(1.7-0)/0.5
delta

# Check the table for alpha = 0.05, degrees of freedom = 48, and delta = 3.4
# For delta = 3.0 and degrees of freedom = 40; power = 0.83
# For delta = 4.0 and degrees of freedom = 5-; power = 0.84
```

## Question 2
```{r setup, include=FALSE}
library(readxl)

### Import CSV file
data1 <- read.csv("/Users/erinanderson/Desktop/R_Work/Stats_HW_1/BMI.csv")

head(data1)
```
## a
```{r 2-a, include=TRUE}
# a) Calculate your confidence interval by hand. Show your work. (i.e., show your calculation for the point estimate, the standard error term, the t-value, and the steps in calculating the confidence interval.)

#HW 1: b0, b1, SSxx, xbar, MSE.

# b1 +/- SE(b1)*(1-alpha/2; n-2)

#SE(b1)

SE <- sqrt(SS_xx/(n-2))

# H0: b1 = 0; Ha: b1 != 0.

### T-test
t_test <- b1/MSE
t_test

```

## b
```{r 2-b, include=TRUE}
#For the BMI.csv data, construct a 90% confidence interval for mean BMI score when weight is 100kg.

# b) Calculate your confidence interval, this time using the predict-function in R. Copy-paste your R commands, and the output showing the interval.

### Regression model
lm1 <- lm(BMI ~ Weight.kg., data1)

#### Create a new data frame and set the value of X
data1.new <- data.frame(Weight.kg.=100)

#### Constructing a confidence interval for the mean response when X=100.
#### Apply the predict function and set the interval type as "confidence"
#### We will use a 0.90 confidence level (default=0.95).
predict(lm1, newdata=data1.new, interval="confidence", level=0.90)
```


## c) Interpret your interval in the context of the problem.

90% of the confidence intervals produced for a person with weight = 100kg capture beta1 and 10% do not. Therefore, we can be 90% confident that Ha is true

