---
title: "HW_2"
author: "Erin Anderson"
date: "2026-09-27"
output: html_document
---
## **Question 1**

```{r 1, include=TRUE}
# The data for this problem is on the speed (X, in mph) and stopping distance (Y, in ft) of 50 cars. What would be the power of an α=.05 test if the true slope of the regression line for this problem is 1.7? Assume σ{b1} = 0.5. Write a sentence explaining what this power means.

# delta = abs(beta1 - beta(10))/sigma(beta1)
# H0: b1 = 0

delta <- abs(1.7-0)/0.5
delta
```
Check the table for alpha = 0.05, degrees of freedom = 48, and delta = 3.4

**For delta = 3.0 and degrees of freedom = 40; power = 0.83**

**For delta = 4.0 and degrees of freedom = 5; power = 0.84**

# **Question 2**

### For the BMI.csv data, construct a 90% confidence interval for mean BMI score when weight is 100kg.
```{r setup, include=FALSE}
library(readxl)

### Import CSV file
data1 <- read.csv("/Users/erinanderson/Desktop/R_Work/Stats_HW_1/BMI.csv")

head(data1)
```
### a
```{r 2-a, include=TRUE}
# a) Calculate your confidence interval by hand. Show your work. (i.e., show your calculation for the point estimate, the standard error term, the t-value, and the steps in calculating the confidence interval.)

#HW 1: b0, b1, SSxx, xbar, MSE.

# y(hat) +/- var(y)*t(star)

b1 = 0.3530631
SS_xx = 275969.7
n = 255
MSE = 57.20585
b0 = 0.07602281
x_bar = 105.698 #mean_x

# Point estimate when X(h) = 100kg
# Y(hat) = b0 + b1*X(h)

y_hat <- b0 + b1*100
y_hat

# Variance estimate
# s^2 (y-hat) = MSE(1/n + ((X(h) - X(bar)))^2/(sum(X(i) - x(bar))))

numerator <- (100 - x_bar)^2
numerator

denominator <- sum((data1$Weight.kg. - x_bar)^2)
denominator

s <- numerator/denominator
s <- (1/n) + s
s <- MSE * s
s <- sqrt(s)

# H0: b1 = 0; Ha: b1 != 0.

### T-test (using t-table): two-tailed t-test at alpha = 0.9 on 255 degrees of freedom: t = 1.645

t_star <- 1.645

# y(hat) +/- var(y)*t(star)

CI_top <- y_hat + (s*t_star)
CI_top

CI_bottom <- y_hat - (s*t_star)
CI_bottom

cbind(CI_bottom, CI_top)

```

### b
```{r 2-b, include=TRUE}
# b) Calculate your confidence interval, this time using the predict-function in R. Copy-paste your R commands, and the output showing the interval.

### Regression model
lm1 <- lm(BMI ~ Weight.kg., data1)

#### Create a new data frame and set the value of X
data1.new <- data.frame(Weight.kg.=100)

#### Constructing a confidence interval for the mean response when X=100.
#### Apply the predict function and set the interval type as "confidence"
#### We will use a 0.90 confidence level (default=0.95).
predict(lm1, newdata=data1.new, interval="confidence", level=0.90)
```


### c) Interpret your interval in the context of the problem.

**90% of the confidence intervals produced for a person with weight = 100kg capture beta1 and 10% do not. Therefore, we can be 90% confident that Ha is true**

# **Question 3**
### For the BMI.csv data, construct a 90% prediction interval for mean BMI score of a person having weight 100kg.
### a

```{r 3-a, include=TRUE}
#a) Calculate your prediction interval by hand. Show your work. (i.e., show your calculation for the point estimate, the standard error term, the t-value, and the steps in calculating the prediction interval.)

# y(hat) +/- var(pred)

b1 = 0.3530631
SS_xx = 275969.7
n = 255
MSE = 57.20585
b0 = 0.07602281
x_bar = 105.698 #mean_x

# Point estimate when X(h) = 100kg
# Y(hat) = b0 + b1*X(h)

y_hat <- b0 + b1*100
y_hat

# Variance estimate
# s^2 (pred) = MSE + s^2(y-hat)

# Var(y_hat)
numerator <- (100 - x_bar)^2
numerator

denominator <- sum((data1$Weight.kg. - x_bar)^2)
denominator

s <- numerator/denominator
s <- (1/n) + s
s <- MSE * s

# Not taking square root until the end
# Var(pred)

s_pred <- s + MSE
s_pred <- sqrt(s_pred)

### T-test (using t-table): two-tailed t-test at alpha = 0.9 on 255 degrees of freedom: t = 1.645

t_star <- 1.645

# y(hat) +/- var(pred)*t(star)

PI_top <- y_hat + (s_pred*t_star)
PI_top

PI_bottom <- y_hat - (s_pred*t_star)
PI_bottom

cbind(PI_bottom, PI_top)
```

### b
```{r 3-b, include=TRUE}
# b) Calculate your prediction interval, this time using the predict-function in R. Copy-paste your R commands, and the output showing the interval.

### Regression model
lm1 <- lm(BMI ~ Weight.kg., data1)

#### Create a new data frame and set the value of X
data1.new <- data.frame(Weight.kg.=100)

#### Constructing a prediction interval for the mean response when X=100.
#### Apply the predict function and set the interval type as "predict"
#### We will use a 0.90 confidence level (default=0.95).
predict(lm1, newdata=data1.new, interval="predict", level=0.90)
```

### c

**With 90% confidence, we predict that the BMI for the next person with weight = 100kg will be somewhere between 22.87 and 47.89**

## **Question 4**

### # For the BMI.csv data, construct a 90% prediction interval for mean BMI score of three persons each of whose weight is 100kg

### a
```{r 4-a, include=TRUE}
# a) Calculate your prediction interval by hand. Show your work. (i.e., show your calculation for the point estimate, the standard error term, the t-value, and the steps in calculating the prediction interval.)

# y(hat) +/- var(mean_pred)

b1 = 0.3530631
n = 255
MSE = 57.20585
b0 = 0.07602281
x_bar = 105.698 #mean_x

# Point estimate when X(h) = 100kg
# Y(hat) = b0 + b1*X(h)

y_hat <- b0 + b1*100
y_hat

# Variance estimate
# s^2 (predmean) = MSE/m + s^2(y-hat)

# Var(y_hat)
numerator <- (100 - x_bar)^2
numerator

denominator <- sum((data1$Weight.kg. - x_bar)^2)
denominator

s <- numerator/denominator
s <- (1/n) + s
s <- MSE * s

# Not taking square root until the end
# Var(predmean)

s_predmean <- s + (MSE/3)
s_predmean <- sqrt(s_predmean)

### T-test (using t-table): two-tailed t-test at alpha = 0.9 on 255 degrees of freedom: t = 1.645

t_star <- 1.645

# y(hat) +/- var(predmean)*t(star)

PI_top_m <- y_hat + (s_pred*t_star)
PI_top_m

PI_bottom_m <- y_hat - (s_pred*t_star)
PI_bottom_m

cbind(PI_bottom_m, PI_top_m)
```

### b

```{r 4-b, include=TRUE}
# b) Calculate your prediction interval, this time using R – you can do so using the following commands. Copy-paste your R commands, and the output showing the interval

library(ALSM)
### lm1 is your linear model
### new.dat is a data frame with the new X value
lm1 <- lm(BMI ~ Weight.kg., data1)
new.dat <- data.frame(X=100)

###For prediction of one new observation:
ci.reg(lm1, new.dat, type = 'n', alpha=0.10,m=1)

###For prediction of mean of 3 new observations
ci.reg(lm1, new.dat, type = 'nm', alpha=0.10,m=3)
```

### c
```{r 4-c, include = TRUE}
#c) Using your prediction interval for mean BMI score of the three persons, calculate the prediction interval for their total BMI. (This is not very informative here, but do it as an exercise.)

#We can get the prediction interval for the total number of work hours for the three people at weight = 100kg by multiplying the above prediction limits by n = 3:

# 3 * (28.12971) = 84.38913
# 3 * (42.63497) = 127.9049
```

**So we can predict, with 90% confidence, that all three people at weight = 100kg will have a combined BMI value between 84.39 and 127.90**

# **Question 5**
### Use R to find and display the 95% confidence band for the simple linear regression line predicting BMI using weight. Create a scatter plot, show the regression line and the confidence band in your plot. Label your axes appropriately. Paste your R-code and output for this answer.

```{r 5, include = TRUE}
# Create linear regression 
lm1 <- lm(BMI ~ Weight.kg., data1)

#### Apply the predict function and set the interval type as "confidence"
pred <- predict(lm1, newdata=data1, interval="confidence", level=0.95)

# Combine the predictions with the original data
pred_df <- cbind(data1, pred)

library(ggplot2)

p <- ggplot(pred_df, aes(x = Weight.kg., y = BMI)) +
  geom_point() +
  geom_ribbon(aes(ymin = lwr, ymax = upr), fill = "blue", alpha = 0.4) +
  geom_line(aes(y = fit)) +
  labs(
    x = "Weight (kg)",
    y = "BMI") +
  theme_minimal()

p

```

